---
title: "Comparison of available dice packages"
author: "Verena Haunschmid"
date: "August 28, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Available packages

Besides `alea` other packages related to casting dice are available. In case you are looking for functionality which is not provided by `alea`, this list gives you an overview over the packages I found and the description they provided: 

* [dice](https://github.com/cran/dice): Calculate probabilities of various dice-rolling events
* [Rdice](https://github.com/gennaro-tedesco/Rdice): R package to simulate dice rolls, combinatorical and best choice problems.
* [diceSyntax](https://github.com/oganm/diceSyntax): dice roller with standard syntax

This document shows a few of the use cases where another package might be more suitable for you.

### dice

```{r dice1}
library(dice)
```

The package `dice` is a package for calculating probabilities of various dice-rolling events, so this is something completely different from what `alea` is doing. The package consists of two functions: (a) `getEventProb()` and (b) `getSumProbs()`. The package is not very well documented and the functionality of `getEventProb()` was not immediately clear to me. So I try to explain it in more detail.

The first three examples are taken from the manual of the package.

```{r dice2}
getEventProb(nrolls = 6,
             ndicePerRoll = 1,
             nsidesPerDie = 4,
             eventList = list(4, 3, c(1,2)),
             orderMatters = FALSE)
```

```{r dice3}
getEventProb(nrolls = 3,
             ndicePerRoll = 2,
             nsidesPerDie = 6,
             eventList = list(10, 4, c(2:6, 8:12)),
             orderMatters = TRUE)
```

The function `getSumProbs()` is more straightforward. 

```{r dice4}
getSumProbs(ndicePerRoll = 5,
            nsidesPerDie = 6,
            nkept = 3,
            dropLowest = TRUE)
```

For better understanding, let me compute the probabilities of a more commonly used example (e.g., for teaching). Casting a 6-sided dice for two times.

```{r dice5}
probs <- getSumProbs(ndicePerRoll = 2,
            nsidesPerDie = 6,
            nkept = 2)
probs <- data.frame(probs$probabilities)
ggplot(probs) + geom_col(aes(x=Sum,  y=Ways.to.Roll)) + scale_x_discrete(limits=probs$Sum)
```

### Rdice

The package `Rdice` contains a set of function that can do similar things as `alea`.

```{r Rdice1}
library(Rdice)
```

The core function seems to be `dice.roll()`. It produced a lot of output in a complicated list structure which does not make it very useful for other applications.

The following example is taken from the help page of the function an casts 3 6-sided dice 5 times each.

```{r Rdice2}
dice.roll(faces = 6, dice = 3, rolls = 5)
```

The package also contains three datasets of non-transitive dice, i.e. [Efron dice](http://mathworld.wolfram.com/EfronsDice.html) (`efron`), [Miwin dice](https://en.wikipedia.org/wiki/Miwin%27s_dice) (`miwin`) and [Oskar](https://en.wikipedia.org/wiki/Nontransitive_dice#Oskar_dice) (`oskar`).

The package contains a [vignette](https://github.com/gennaro-tedesco/Rdice/blob/master/vignettes/Rdice-vignette.pdf) with more detailed explanations.

### diceSyntax

The package `diceSyntax` contains several functions but only documentation for `roll`.

The package is only available on github.

```{r diceSyntax1, eval=FALSE}
devtools::install_github("oganm/diceSyntax")
```

```{r diceSyntax2}
library(diceSyntax)
```

The interesting part of this package is that you can pass dice events like `4d6` in order to cast 4 6-sided dice.

```{r diceSyntax3}
roll("4d6")
```

More examples (e.g., keep highest/lowest 3, ...) can be found in the [README of diceSyntax](https://github.com/oganm/diceSyntax).
